---
title: "ALL analysis and p-values with different statistical tests"
output: html_document
date: "2023-03-07"
output:
  word_document: default
  html_document: default
---

Install BioCManager and the ALL package

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A data frame containing the expression values for the B-cell ALL patients in stage B, B1, B2, B3, B4 from the ALL data created. 

```{r}
library(ALL)
data(ALL)
ALL.bcell<-ALL[,ALL$BT %in% c("B","B1","B2","B3","B4")]
ALL.bcell
```

Number of patients per group

```{r}
table(ALL.bcell$BT)
```

Shapiro test to test normality and p-values collected, 

```{r}
ALL.sw.pvalues<-apply(exprs(ALL.bcell),1,function(x) shapiro.test(residuals(lm(x~ALL.bcell$BT)))$p.value)
head(ALL.sw.pvalues)
```

Homoscedasticity assumption performed

```{r}
library(lmtest)
ALL.bp.pvalues<-apply(exprs(ALL.bcell),1,function(x) as.numeric(bptest(lm(x~ALL.bcell$BT),studentize=FALSE)$p.value))
head(ALL.bp.pvalues)
```

Counts for each and both 


```{r}
sum(ALL.sw.pvalues>0.05)
```

```{r}
sum(ALL.bp.pvalues>0.05)
```

```{r}
sum(ALL.sw.pvalues>0.05 & ALL.bp.pvalues>0.05)
```
p-values from ANOVA collected
```{r}
ALL.anova.pvalues<-apply(exprs(ALL.bcell),1,function(x) anova(lm(x~ALL.bcell$BT))$Pr[1])
head(ALL.anova.pvalues)
```

ANOVA p-values less than 0.000001 

```{r}
featureNames(ALL.bcell)[ALL.anova.pvalues<0.00001]
```

Kruskal wallis p-values collected

```{r}
ALL.kw.pvalues<-apply(exprs(ALL.bcell),1,function(x) kruskal.test(x~ALL.bcell$BT)$p.value)
head(ALL.kw.pvalues)
```

Kw p values less than 0.000001 

```{r}
featureNames(ALL.bcell)[ALL.kw.pvalues<0.000001]
```

Comparing ANOVA and KW p-values to less than 0.001

```{r}
small.anova<-ALL.anova.pvalues<0.001
small.kw<-ALL.kw.pvalues<0.001
table(small.anova,small.kw)
```
291 genes have pvalue smaller than 0.001 for both ANOVA and Kruskal-Wallis.

124 genes have pvalue smaller than 0.001 for ANOVA.

38 genes have pvalue smaller than 0.001 for Kruskal-Wallis.